---
title: "NBA - Last three seasons analysis"
author: "Juanma Martinez"
date: "10/05/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
NBAstas_Tableau_right <- read_csv("NBAstas_Tableau-right.csv")
View(NBAstas_Tableau_right)

colnames(NBAstas_Tableau_right)

library(dplyr)


# Eastern Conference ------------------------------------------------------


NBA_data_1<-NBAstas_Tableau_right %>% filter(Season_1 == "2018-2019" & Conference == "Eastern Conference") %>%
  select(1,4,5,6)
View(NBA_data_1)

names(NBA_data_1)[1]<-"Team"
names(NBA_data_1)[2]<-"Wins_1"
names(NBA_data_1)[3]<-"Losses_1"
names(NBA_data_1)[4]<-"Win_%_1"

NBA_data_2<-NBAstas_Tableau_right %>% filter(Season_1 == "2017-2018" & Conference == "Eastern Conference") %>%
  select(1,4,5,6)


names(NBA_data_2)[1]<-"Team"
names(NBA_data_2)[2]<-"Wins_2"
names(NBA_data_2)[3]<-"Losses_2"
names(NBA_data_2)[4]<-"Win_%_2"

View(NBA_data_2)



d <-inner_join(NBA_data_1,NBA_data_2, by = ("Team"))

View(d)



NBA_data_3<-NBAstas_Tableau_right %>% filter(Season_1 =="2016-2017" & Conference == "Eastern Conference") %>%
  select(1,4,5,6)

names(NBA_data_3)[1]<-"Team"
names(NBA_data_3)[2]<-"Wins_3"
names(NBA_data_3)[3]<-"Losses_3"
names(NBA_data_3)[4]<-"Win_%_3"

View(NBA_data_3)

d_2 <-inner_join(d,NBA_data_3, by = ("Team"))
View(d_2)





# Western Conference ------------------------------------------------------



NBA_data_A<-NBAstas_Tableau_right %>% filter(Season_1 == "2018-2019" & Conference == "Western Conference") %>%
  select(1,4,5,6)
View(NBA_data_A)

names(NBA_data_A)[1]<-"Team"
names(NBA_data_A)[2]<-"Wins_1"
names(NBA_data_A)[3]<-"Losses_1"
names(NBA_data_A)[4]<-"Win_%_1"

NBA_data_B<-NBAstas_Tableau_right %>% filter(Season_1 == "2017-2018" & Conference == "Western Conference") %>%
  select(1,4,5,6)


names(NBA_data_B)[1]<-"Team"
names(NBA_data_B)[2]<-"Wins_2"
names(NBA_data_B)[3]<-"Losses_2"
names(NBA_data_B)[4]<-"Win_%_2"

View(NBA_data_B)



d_A_B <-inner_join(NBA_data_A,NBA_data_B, by = ("Team"))

View(d_A_B)


NBA_data_C<-NBAstas_Tableau_right %>% filter(Season_1 =="2016-2017" & Conference == "Western Conference") %>%
  select(1,4,5,6)

names(NBA_data_C)[1]<-"Team"
names(NBA_data_C)[2]<-"Wins_3"
names(NBA_data_C)[3]<-"Losses_3"
names(NBA_data_C)[4]<-"Win_%_3"

View(NBA_data_C)

d_A_B_C <-inner_join(d_A_B,NBA_data_C, by = ("Team"))
View(d_2)



library(gt)



```

## NBA Regular season analysis - Eastern conference


NBA Eastern conference stats for the last three seasons. Regular seasons stats (Wins, Losses , Percentage of Wins).




```{r cars, echo=FALSE}
NBA_gt_eastern<- d_2 %>% gt(
  rowname_col = "Team",
)%>%
  tab_header(
    title = md("**NBA** Stats"),
    subtitle = md("Regular Season - Eastern Conference")
  ) %>% 
  fmt_percent(
    columns = vars(`Win_%_1`,`Win_%_2`,`Win_%_3`),
    decimals = 1
  ) %>%
  tab_footnote(
    footnote = "Regular Season Analysis",
    locations = cells_title(groups = "subtitle")
  ) %>%
  tab_source_note(
    source_note = md(
      "Source: NBA official page [stats.nba.com](https://stats.nba.com).")
  )  %>%
  tab_stubhead_label(label = md("**TEAM**")) %>%
  tab_options(
    table.width = pct(100),
    #table.background.color = "lightgray",
    row.padding = px(5),
    heading.title.font.size = "small",
    heading.subtitle.font.size = "small",
    heading.background.color = "red",
    column_labels.background.color = "lightgray"
  ) %>%
  tab_spanner(
    label = md("2018-2019"),
    columns = vars(`Wins_1`,`Losses_1`,`Win_%_1`)
  ) %>%
  tab_spanner(
    label = md("2017-2018"),
    columns = vars(`Wins_2`,`Losses_2`,`Win_%_2`)
  ) %>%
  tab_spanner(
    label = md("2016-2017"),
    columns = vars(`Wins_3`,`Losses_3`,`Win_%_3`)
  ) %>%
  data_color(
    columns = vars(`Win_%_1`,`Win_%_2`,`Win_%_3`),
    colors = scales::col_numeric(
      palette = c(
        "lightgray", "lightgreen", "lightblue"),
      domain = c(0.2, 0.5,0.74))
  ) %>%
  cols_label(
    Wins_1 = md("Wins"),
    Losses_1 = md("Losses"),
    `Win_%_1`= md("Wins %"),
    Wins_2 = md("Wins"),
    Losses_2 = md("Losses"),
    `Win_%_2`= md("Wins %"),
    Wins_3 = md("Wins"),
    Losses_3 = md("Losses"),
    `Win_%_3`= md("Wins %")
  )

NBA_gt_eastern

```

## NBA Regular season analysis - Western conference


NBA Western conference stats for the last three seasons. Regular seasons stats (Wins, Losses , Percentage of Wins).


```{r pressure, echo=FALSE}
NBA_gt_Western<- d_A_B_C  %>% gt(
  rowname_col = "Team",
)%>%
  tab_header(
    title = md("**NBA** Stats"),
    subtitle = md("Regular Season - Western Conference")
  ) %>% 
  fmt_percent(
    columns = vars(`Win_%_1`,`Win_%_2`,`Win_%_3`),
    decimals = 1
  ) %>%
  tab_footnote(
    footnote = "Regular Season Analysis",
    locations = cells_title(groups = "subtitle")
  ) %>%
  tab_source_note(
    source_note = md(
      "Source: NBA official page [stats.nba.com](https://stats.nba.com).")
  )  %>%
  tab_stubhead_label(label = md("**TEAM**")) %>%
  tab_options(
    table.width = pct(100),
    #table.background.color = "lightgray",
    row.padding = px(5),
    heading.title.font.size = "small",
    heading.subtitle.font.size = "small",
    heading.background.color = "red",
    column_labels.background.color = "lightgray"
  ) %>%
  tab_spanner(
    label = md("2018-2019"),
    columns = vars(`Wins_1`,`Losses_1`,`Win_%_1`)
  ) %>%
  tab_spanner(
    label = md("2017-2018"),
    columns = vars(`Wins_2`,`Losses_2`,`Win_%_2`)
  ) %>%
  tab_spanner(
    label = md("2016-2017"),
    columns = vars(`Wins_3`,`Losses_3`,`Win_%_3`)
  ) %>%
  data_color(
    columns = vars(`Win_%_1`,`Win_%_2`,`Win_%_3`),
    colors = scales::col_numeric(
      palette = c(
        "lightgray", "lightgreen", "lightblue"),
      domain = c(0.2, 0.5,0.82))
  ) %>%
  cols_label(
    Wins_1 = md("Wins"),
    Losses_1 = md("Losses"),
    `Win_%_1`= md("Wins %"),
    Wins_2 = md("Wins"),
    Losses_2 = md("Losses"),
    `Win_%_2`= md("Wins %"),
    Wins_3 = md("Wins"),
    Losses_3 = md("Losses"),
    `Win_%_3`= md("Wins %")
  )


NBA_gt_Western
```


**R Packages used**


- readr
- dplyr
- gt



**Code, questions and feedback**

The code for this article is published in my [github account](https://github.com/JuanmaMN/gt-package).  


Please feel free to reach out for any question or feedback.
